---
title: "R Notebook"
output: html_notebook
---

# load libs
```{r, echo=FALSE, include=FALSE}
library(tidyr)
library (dplyr)
library (ggplot2)
library(agricolae)
library(knitr)
library(lubridate)
library(GGally)
```

## 1. Maize

# Read and check maize biomass
```{r}
maize_bio_raw <- 
  read.table("K:\\Rainshelter\\Maize 2012-13\\Analysis\\WorkedData\\Biomass\\BiomassForR.txt", header=TRUE)

str(maize_bio_raw)

# define N treat categories
# n_cat_maize <- data.frame(NitroTreat = c("000N","075N","250N"), 
#                           NitroCat = c("Low","Median","High"))

# clean up and massage
maize_bio_work <- maize_bio_raw %>%
  mutate(HarvestDate = dmy(Date)) %>%
  mutate(Block = as.factor(Block),
         Trt = as.factor(Trt),
         Plot = as.factor(Plot)
         ) %>%
  dplyr::select(Date, Plot, WaterTreat, NitroTreat, Grain, Leaf, Stem, Rachis, Dead) %>%
  mutate(HarvestDate = dmy(Date), 
       #  Total = (Leaf + Stem + Dead + Grain + Rachis),
         Nitrogen =  as.numeric(as.character(gsub("N","",NitroTreat))),
         NitroTreat = factor(NitroTreat,levels = c("000N","075N","250N"), 
                             labels= c("Low N","Median N","High N")),
         WaterTreat = ifelse(WaterTreat=="Irrig", "Irrigated","Dryland"), # same as wheat
         WaterTreat = as.factor(WaterTreat) ,
         Irrigation = as.factor(WaterTreat)) %>%
         tidyr::gather("Organ","BiomassWeight",Grain:Dead) %>%
         mutate(Organ = ifelse(Organ == "Dead", "Senesced", as.character(Organ))) %>%
         mutate(Organ = as.factor(Organ))


# function to do percentages
# ppf <- function(organ,total) {
#   
#   round(organ/total*100,1)
#   
# }
# mutate(
#     Total = (Leaf + Stem + Dead + Grain + Rachis),
#        #  TestTotal = round(Total/Total*100,0),
#          LeafPerc = ppf(Leaf,Total),
#          StemPerc = ppf(Stem,Total),
#          DeadPerc = ppf(Dead,Total),
#          GrainPerc = ppf(Grain,Total),
#          RachisPerc = ppf(Rachis,Total)
#                              ) 


# FIXME: gather organ names in a "Organ" column


#head(maize_bio_work)
summary (maize_bio_work)
#str(maize_bio_work)
```

# Read and check maize nitrogen
```{r}
maize_N_raw <- read.table("K:\\Rainshelter\\Maize 2012-13\\Analysis\\WorkedData\\Nitrogen\\plantN.txt", header = TRUE)
#duplicated(maize_N_raw)

str(maize_N_raw)

# clean up and massage
maize_N_work <- maize_N_raw %>%
  mutate(HarvestDate = dmy(Date)) %>%
  mutate(Block = as.factor(Block),
         Plot = as.factor(Plot),
         HarvestDate = dmy(Date), 
         Organ = Component,
         Nitrogen =  as.numeric(as.character(gsub("N","",NitroTreat))),
         Irrigation = as.factor(WaterTreat),
         Organ = ifelse(Organ == "Dead", "Senesced", as.character(Organ)), # make names align with N data
         Organ = ifelse(Organ == "Chaff", "Rachis", as.character(Organ)),
         NitrogenPerc = Nitrogen_Perc , # to match wheat
         Organ = as.factor(Organ)
         )

str(maize_N_work)
head(maize_N_work)
summary (maize_N_work)
```

# merge biomass and N in maize

```{r}

# revove duplicated fields before merge
maize_N_work$WaterTreat <- NULL
maize_N_work$NitroTreat <- NULL
maize_N_work$Nitrogen <- NULL
maize_N_work$Irrigation <- NULL
#maize_N_work$Block <- NULL
maize_N_work$Component <- NULL
maize_N_work$Date <- NULL

# check fields
colnames(maize_bio_work)
colnames(maize_N_work)

# merge the baby
maize_bio_N_work <- merge(maize_bio_work, maize_N_work, by=c("HarvestDate","Plot","Organ"))

# check
maize_bio_N_work$NitrogenWeight <- 
  maize_bio_N_work$BiomassWeight * maize_bio_N_work$Nitrogen_Perc/100

summary(maize_bio_N_work)


```

# overall check
```{r}
maize_bio_N_work %>%
dplyr::select(Organ,BiomassWeight:NitrogenWeight) %>%
ggpairs(aes(colour=Organ))
  
```

# find odd points
```{r}
# graph N amounts from raw data and from R calculations
maize_bio_N_work %>%
  ggplot(aes(x=NitrogenWeight, y=N_amount, colour=NitroTreat, shape=Irrigation)) +
  geom_point() +
  geom_abline(intercept=0,slope=1)

# FIXME: identify points that diverge from x=y

```

## 2) Wheat

# Read and check wheat biomass

Data from exel sheet at  K:\Rainshelter\Wheat 2015-16\Analysis\ArbitratorTestData
```{r}

wheat_bio_raw <- 
  read.table("K:\\Rainshelter\\Wheat 2015-16\\Analysis\\Biomass\\BiomassLAIData.txt", header=TRUE)

#str(wheat_bio_raw)

head(wheat_bio_raw)

```

```{r}
# clean up and massage
wheat_bio_work <- wheat_bio_raw %>%
  dplyr::select(HarvestDate,Plot,
                Block, Nitrogen,Irrigation,
                Leaf:Chaff) %>%
  mutate(HarvestDate = dmy(HarvestDate)) %>%
  mutate(Block = as.factor(Block),
      #   Trt = as.factor(TreatNo),
         Plot = as.factor(Plot-100),
         WaterTreat = Irrigation,
         NitroTreat = factor(Nitrogen,levels = c(0,50,250), labels= c("Low N","Median N","High N"))) %>% # Nil,Mid,High
  dplyr::select(HarvestDate, Plot,Block, WaterTreat, NitroTreat,Leaf:Chaff) %>%
   tidyr::gather("Organ","BiomassWeight",Leaf:Chaff) %>%
    mutate(Organ = factor(Organ, levels=c("Leaf",	"Stem",	"Dead",	"Ear",	"Grain",	"Chaff"), 
                          labels = c("Leaf",	"Stem",	"Senesced",	"Ear",	"Grain",	"Rachis"))) %>%
   filter(Organ != "Ear") # in raw data now loaded into "Chaff" or now "Rachis"

# FIXME: PLot ear and Chaff or Rachis to be sure all ok

summary(wheat_bio_work)
```

# Read wheat nitrogen

```{r}
wheat_N_raw <- 
  read.table("K:\\Rainshelter\\Wheat 2015-16\\Analysis\\N_Plant\\NitrogenRawWheat.txt", header = TRUE)
#duplicated(maize_N_raw)

str(wheat_N_raw)

# clean up and massage
wheat_N_work <- wheat_N_raw %>%
  mutate(HarvestDate = dmy(Date)) %>%
  mutate(Block = as.factor(Block),
         Plot = as.factor(as.numeric(Plot-100)),
         HarvestDate = dmy(Date), 
         Nitrogen =  as.numeric(Nitrogen),
         Irrigation = as.factor(Irrigation)
         ) %>%
      mutate(Organ = factor(Organ, levels=c("Leaf",	"Stem",	"Dead",	"Ear",	"Grain",	"Rachis"), 
                          labels = c("Leaf",	"Stem",	"Senesced",	"Ear",	"Grain",	"Rachis"))) 

str(wheat_N_work)
head(wheat_N_work)
summary (wheat_N_work)
```

# merge wheat biomass and N

```{r}
# revove duplicated fields before merge
wheat_N_work$Block <- NULL

# check fields
summary(wheat_bio_work)
summary(wheat_N_work)

```

```{r}
# merge the baby
wheat_bio_N_work <- merge(wheat_bio_work, wheat_N_work, by=c("HarvestDate","Plot","Organ"))

# add N amount field
wheat_bio_N_work$NitrogenWeight <- wheat_bio_N_work$BiomassWeight * wheat_bio_N_work$NitrogenPerc/100
  
  
summary(wheat_bio_N_work)
```

## 3. Merge maize and wheat

# Check and clean both wheat and maize
```{r}
print("====================================================")
print("========        MAIZE                   ============")
print("====================================================")

maize_bio_N_work$Date <- maize_bio_N_work$Nitrogen_Perc <- maize_bio_N_work$N_amount <-NULL
maize_bio_N_work$Crop <- "Maize"
summary(maize_bio_N_work)


print("====================================================")
print("========        WHEAT                   ============")
print("====================================================")

wheat_bio_N_work$Date <- wheat_bio_N_work$Partition <- 
wheat_bio_N_work$TrtNo <- wheat_bio_N_work$CarbonPerc <-NULL
wheat_bio_N_work$Crop <- "Wheat"
summary(wheat_bio_N_work)
```

# Bind wheat and maize
```{r}
maize_wheat_bio_N_work <- rbind(maize_bio_N_work, wheat_bio_N_work)

maize_wheat_bio_N_work$Crop <- as.factor(maize_wheat_bio_N_work$Crop)

summary(maize_wheat_bio_N_work)

```

