---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---

# load libs
```{r, echo=FALSE, include=FALSE}
library(tidyr)
library (dplyr)
library (ggplot2)
library(agricolae)
library(knitr)
library(lubridate)
library(GGally)
```


#  RUE calculation

## Merge Biomass and Interception

```{r}

bio_raw <- read.csv("C:\\GitHubRepos\\Arbitrator\\Analysis\\Biomass_N_merged_WheatMaize.csv")

summary(bio_raw)

```

```{r}
# clean data
bio_work <- bio_raw %>%
  na.omit() %>%
  dplyr::select(-X,-NitrogenPerc,-NitrogenWeight) %>%
  mutate(Block = as.factor(Block),
         HarvestDate = ymd(HarvestDate),
         WaterTreat = as.factor(WaterTreat) ,
         WaterTreat = factor(WaterTreat, 
                             levels= c("Dryland","Irrigated")),
         NitroTreat = as.factor(NitroTreat),
         NitroTreat = factor(NitroTreat, 
                             levels= c("Low N","Median N","High N"))) %>%
  dplyr::select(HarvestDate, Crop, Organ, NitroTreat, WaterTreat, Block,BiomassWeight) %>%
  tidyr::spread(Organ,BiomassWeight) %>%
  rowwise() %>%
  mutate(TotalBiomass = sum(Grain,Leaf,Rachis,Senesced,Stem, na.rm=TRUE)) %>%
  dplyr::select(-(Grain:Stem))

summary(bio_work)
```
# check
```{r, warning=FALSE}

# how many onservations in each?
bio_work %>%
group_by(Crop,NitroTreat,WaterTreat,Block) %>%
 summarise(n = n())


```

# retrieve intercepted radiation

- accumulated radiation per plot

```{r}

rad_sum_df <- read.csv("C:\\GitHubRepos\\Arbitrator\\Analysis\\RadIntSum.csv") %>%
  dplyr::select(-X) # FIXME: clean this early

summary(rad_sum_df)

```



# Merge biomass and Sum intercepted radiation

```{r}

buf1 <- rad_sum_df %>% dplyr::select(-radn,-RadInt)

bio_rad <- merge(bio_work, buf1, by = c("HarvestDate","Crop","NitroTreat","WaterTreat","Block")) %>%
  mutate(HarvestDate=ymd(HarvestDate))

buf1 <- NULL

summary(bio_rad)
```

# plot and check it

- Biomass as function of accumulated radiation interception

```{r,fig.width=12, fig.height=8}
bio_rad %>%
  mutate(NitroTreat = factor(NitroTreat, 
                             levels= c("Low N","Median N","High N"))) %>%
  ggplot(aes(x=RadIntSum, y=TotalBiomass, colour=WaterTreat, shape=NitroTreat)) +
  geom_point() +
  geom_line(aes(linetype=NitroTreat)) +
# geom_smooth(method="lm", aes(linetype=NitroTreat,fill=WaterTreat), alpha=0.2) +
 scale_linetype_manual(values = c("dotted", "dashed", "solid","dotted", "dashed", "solid"))+
  facet_wrap(Crop~WaterTreat+Block, scales="free", ncol=4)
```

# Define data points suitable for RUE calculation

- Remove points when net increase in shoot biomass was not linear
- e.g. sowing date and last harvest when senescence was high

```{r}
# maximum biomass observed in that plot
bio_X <- bio_rad %>%
  dplyr::select(Crop, NitroTreat, WaterTreat, Block,TotalBiomass, RadIntSum) %>%
  group_by(Crop, NitroTreat, WaterTreat, Block) %>%
  mutate(bio_X = max(TotalBiomass), RadAtBioMax = RadIntSum) %>%
  rowwise() %>%
  filter(TotalBiomass == bio_X) %>%
  dplyr::select(-RadIntSum,-TotalBiomass)

summary(bio_X)

# buffer to hold value of maxium at plot level
buf2 <-merge(bio_rad,bio_X,by=c("Crop", "NitroTreat", "WaterTreat", "Block"))


bio_rad_filtered <- buf2 %>%
mutate(plotCode = paste0(Crop,"_",
                         NitroTreat,"_" ,
                         WaterTreat,"_",
                         Block)) %>%
 mutate(plotCode = factor(plotCode)) %>%
  group_by(Crop, NitroTreat, WaterTreat, Block) %>%
  rowwise() %>%
  ungroup() %>%
  arrange(RadIntSum) %>%
  filter(TotalBiomass <= bio_X & RadIntSum<RadAtBioMax) 



```

FIXME: logic does not work 100% check


```{r,fig.width=12, fig.height=8}
# graph
bio_rad_filtered %>%
  mutate(NitroTreat = factor(NitroTreat, 
                             levels= c("Low N","Median N","High N"))) %>%
  ggplot(aes(x=RadIntSum, y=TotalBiomass, colour=WaterTreat, shape=NitroTreat)) +
  geom_point() +
  geom_line(aes(linetype=NitroTreat)) +
# geom_smooth(method="lm", aes(linetype=NitroTreat,fill=WaterTreat), alpha=0.2) +
 scale_linetype_manual(values = c("dotted", "dashed", "solid","dotted", "dashed", "solid"))+
  facet_wrap(Crop~WaterTreat+Block, scales="free", ncol=4)
```

```{r}
summary(bio_rad_filtered)
```

```{r,fig.width=12, fig.height=8}
# graph filtered data
bio_rad_filtered %>%
  mutate(NitroTreat = factor(NitroTreat, 
                             levels= c("Low N","Median N","High N"))) %>%
  ggplot(aes(x=RadIntSum, y=TotalBiomass, colour=WaterTreat, shape=NitroTreat)) +
  geom_point() +
  geom_line(aes(linetype=NitroTreat)) +
# geom_smooth(method="lm", aes(linetype=NitroTreat,fill=WaterTreat), alpha=0.2) +
 scale_linetype_manual(values = c("dotted", "dashed", "solid","dotted", "dashed", "solid"))+
  facet_wrap(Crop~Block, scales="free", ncol=4)
```

# calculate RUE at plot level

```{r}

rounds <- unique(bio_rad_filtered$plotCode)

# loop through all the individual periods/treats/experiments
resTable <- data.frame(NULL)

for(i in 1:length(rounds)) {
  
 thisStat <- NULL
  
 buf <- bio_rad_filtered[bio_rad_filtered$plotCode == rounds[i],]
 
 bio <- buf$TotalBiomass/10 
 interc  <- buf$RadIntSum
  
  n_s <- length(bio)
  n_m <- length(interc)
  model <- lm(bio~interc)
  bio_sq <- sum((bio - mean(bio))^2)
  int_sq <- sum((interc - mean(interc))^2)
  r2 <- summary(model)$r.squared
  slope <- model$coefficients[[2]]
  slope_sig <- summary(model)$coefficients[8]
  int <- model$coefficients[[1]]
  int_sig <- summary(model)$coefficients[7]
  bio_max <- max(bio)
  light_max <- max(interc)  
  
 thisStat <- data.frame(Crop = buf$Crop[1], 
                        N_treat = buf$NitroTreat[1], 
                        W_treat = buf$WaterTreat[1],
                        Block = buf$Block[1],
                        n = n_s,
                        RUE = slope, 
                        slope_dif_zero = round(slope_sig,digits=3),
                        intercept=int,
                        int_dif_zero = int_sig,
                        bio_max = bio_max,
                        light_max = light_max,
                        R2=r2)
 
 if(i==1) {
   resTable <- thisStat
 } else {
   resTable <- rbind(resTable,thisStat)
 }
}

resTable$Block <- as.factor(resTable$Block)
summary(resTable)


```

# check poor R2 fits
```{r}
g1 <- resTable %>%
  filter(R2<0.85)
```

# calculate RUE averages

```{r,fig.width=12, fig.height=8}

n_lev_m <- c(0,75,250)
n_lev_w <- c(0,50,250)

resTable %>%
  mutate(N_treat_v = ifelse(Crop == "Maize", 
                           as.character(factor(N_treat, labels=n_lev_m)),
                            as.character(factor(N_treat, labels=n_lev_w)))) %>%
  mutate(N_treat_v=as.numeric(as.character(N_treat_v))) %>%
  group_by(Crop, W_treat,N_treat_v) %>%
  summarise(RUE_mean = mean(RUE), RUE_sd=sd(RUE)) %>%
  ggplot(aes(x=N_treat_v, y=RUE_mean, colour=W_treat)) +
  geom_point(size=2) +
  geom_line(aes(linetype=W_treat))+
  facet_grid(.~Crop) +
  ylab("Radiation use efficiency (g/m2)") +
  xlab("Nitrogen fertiliser rate (kg N/m2)") +
  geom_errorbar(aes(ymin=RUE_mean-RUE_sd/2,
                    ymax=RUE_mean+RUE_sd/2), width = 0.5) +
  facet_grid(.~Crop, scales="free")
  
# FIXME: Has to decide which points get in RUE calculation!
# At which level this should be defined (automated or by eye?)

```
 
 
## get RUE reduction fators for water and N

```{r}

summary(resTable)

# get base value as High N Irrigated plots
RUEpot_byBlock <- resTable %>%
  group_by(Crop,Block) %>%
  mutate(RUEpot = max(RUE)) %>%
  dplyr::select(Crop,Block,RUEpot)

summary(RUEpot_byBlock)

# merge it in original df
resTable_wBase <- merge(resTable, RUEpot_byBlock, by=c("Crop","Block"))
summary(resTable_wBase)




  
  
```
## Potential RUE per crop

- maximum value measured

```{r}
resTable_wBase %>%
  group_by(Crop) %>%
  summarise(av=round(mean(RUEpot),1),sd=round(sd(RUEpot),2), CV=round((sd/av*100),2))
```

```{r,fig.width=12, fig.height=8}
resTable_wBase %>% 
  dplyr::select(Crop,W_treat,N_treat,Block,RUE,RUEpot) %>%
 # mutate(N_treat_n = factor(N_treat,levels=c("Low N","Median N","High N"), labels=c(0,50,250))) %>%
 # mutate(N_treat_n = as.numeric(as.character(N_treat_n)))%>%
  rowwise() %>%
  mutate(fst = RUE/RUEpot) %>%
 # filter(N_treat=="Low N" & Block=="1" & Crop=="Maize")%>%
 # spread(W_treat, fst)
  ggplot(aes(x=N_treat,y=fst, colour=W_treat, shape=N_treat))  +
  geom_point() +
  geom_line(aes(linetype=W_treat))+
  facet_grid(Block~Crop)
```

